Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-29 paper/casestudies.tex

Intuitively, adding structure information to a random generator should provide
better results when verifying a system using random testing.
%
% For instance, by generating values that are no only syntactically but also
% semantically meaningful we would expect to reach deeper stages of our program
% pipeline other than merely input parsing/validation.
%
This section describes two case studies showing that considering structure
information when deriving generators consistenly produces better testing
results.

Instead of restricting our scope to Haskell, in this work we followed a broader
evaluation approach taken previously to compare the performance of other
state-of-the-art derivation techniques
\cite{grieco2017,DBLP:conf/haskell/MistaRH18}.
%
Concretely, we evaluate how including additional structure information when
generating a set of random test cases (often referred as a \emph{corpus})
affects the code coverage obtained when using them to test a given target
program.
%
For this purpose, we targeted two external programs which expect highly
structured inputs, namely \emph{GNU CLISP}---the GNU Common Lisp compiler, and
\emph{HTML Tidy}---a well known HTML refactoring and correction utility.
%
We remark that these applications are not written in Haskell.
%
However, there exist Haskell libraries defining ADTs encoding their input
structure, i.e., Lisp and HTML values respectively. These libraries are:
\emph{hs-zuramaru}, implementing an embedded Lisp intepreter for a small subset
of this programming language, and \emph{html}, defining a combinator library for
constructing HTML values.
%
These libraries also come with serialization functions to map Haskell values
into corresponding test case files.

We started by compiling instrumented versions of the target programs in a way
that they also return the execution path followed in the source code every time
we run them with a given input test case.
%
This let us distinguish the amount of different execution paths that a randomly
generated corpus can trigger.
%
On the other side, we used the ADTs defined on the chosen libraries to derive
random generators using \dragen and \dragenp, including structure information
extracted from the library's codebase in the case of the latter.
%
Then, we evaluated the code coverage triggered by independent, randomly
generated corpora of different sizes varying from 100 to 1000 test cases each.
%
In order to remove any external bias, we derived random generators optimized to
follow a uniform distribution of constructors (and pattern matchings or function
calls in the case \dragenp), and carefully adjusted their generation sizes to
match the average test case size in bytes.
%
This way, any noticeable difference in the code coverage can be attributed to
the present (or lack thereof) structure information when generating the test
cases.
%
Additionally, to achieve statistical significance we repeated each experiment 30
times with independently generated sets of random test cases.

\begin{figure*}[t]
  \centering
  \input{tikz/lisp.tex}
  \hspace{5pt}%
  \input{tikz/html.tex}
  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
    \dragenp (\ref{exp:dragenp}). }
  \label{fig:coverage}
\end{figure*}

Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
triggered for each permutation of corpus size and derivation tool, including
error bars indicating the standard error of the mean on each case.
%
We proceed to describe each case study in detail.

\paragraph{Generating Lisp code using pattern matching information}

In this first case study we wanted to evaluate the observed code coverage
differences when considering structure information present on functions pattern
matchings.

Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
represented as plain strings; and literal values like booleans or integers.
%
In order to interpret Lisp programs, this unified representation of data and
code requires this library to pattern match against common patterns like
let-bindings, if-then-else expressions and arithmetic operators among others.
%
In particular, each one of these patterns match a against special symbol of the
Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
sub-expressions.

In this case study we extracted the structure information from the pattern
matchings present on this library and included it into the generation
specification of our random Lisp values, which were generated by randomly
picking from a total of 6 data constructors and 8 different pattern matchings.
%
By doing this we obtained a consistent code coverage improvement of
approximately $4\%$ using \dragenp with respect to the one obtained with
\dragen (see Figure \ref{fig:coverage} (a)).
%
\todo[inline, author=AM]{Shall we say anything justifing this little
  improvement?}

\paragraph{Generating HTML pages using abstract interface information}

@@

The library we used for this purpose represents HTML values very much in the
same way as we exemplify in Section \ref{sec:randomtesting}, i.e., defining a
small set of general constructions representing plain text and tags---altough
this library also supports HTML tag attributes as well.
%
Then, this representation is extended with a large abstract interface consisting
of combinators representing common HTML tags and tag attributes---equivalent to
the combinators |div|, |bold| and |hr| illustrated before.

In this case study we included the structure information present on the abstract
interface of this library into the generation specification of random HTML
values, resulting in a generation process that randomly picked among 4 data
constructors and 163 abstract combinators.
%
With this large amount of additional structure information, we observed an
increase of up to $83\%$ in the code coverage obtained with \dragenp with
respect to the one observed with \dragen (see Figure \ref{fig:coverage} (b)).

A manual inspection of the corpora generated with each tool revealed us that, in
general terms, the test cases generated with \dragen rarely represent
syntactically correct HTML values, consisting to a large extent of random
strings within and between HTML tag delimeters (|"<"|, |">"| and |"/>"|).
%
On the other hand, test cases generated with \dragenp encode much more
interesting structural information, being mostly syntactically correct.
%
We found that, in many cases, the test cases generated with \dragenp were
parsed, analysed and reported as valid HTML values by the target application.

With these results we are confident that including the structural information
present on the user codebase improves the testing performance.
%
We consider that our approach is particularly useful when the data types
encoding the shape of our data are vague or not sufficiently structured to be
used to derive powerful random generators with the common type-driven
derivation techniques.

>>>>

% Intuitively, adding structure information to a random generator should provide
% better results when verifying a system using random testing.
%
% For instance, by generating values that are no only syntactically but also
% semantically meaningful we would expect to reach deeper stages of our program
% pipeline other than merely input parsing/validation.
%
This section describes two case studies showing that considering additional
structural information when deriving generators can consistently produce better
testing results in terms of code coverage.
%
Instead of restricting our scope to Haskell, in this work we follow a broader
evaluation approach taken previously to compare state-of-the-art techniques to
derive random data generators based on ADT definitions
\cite{grieco2017,DBLP:conf/haskell/MistaRH18}.
%

We evaluate how including additional structural information when generating a
set of random test cases (often referred as a \emph{corpus}) affects the code
coverage obtained when testing a given target program.
%
For that, we considered two external programs which expect highly structured
inputs, namely \emph{GNU CLISP}---the GNU Common Lisp
compiler\footnote{\href{https://www.gnu.org/software/gcl/}{https://www.gnu.org/software/gcl/}},
and \emph{HTML Tidy}
\footnote{\href{http://www.html-tidy.org}{http://www.html-tidy.org}}---a well
known HTML refactoring and correction utility.
%
We remark that these applications are not written in Haskell.
%
However, there exist Haskell libraries defining ADTs encoding their input
structure, i.e., Lisp and HTML values respectively. These libraries are:
\emph{hs-zuramaru}\footnote{\href{http://hackage.haskell.org/package/zuramaru}{http://hackage.haskell.org/package/zuramaru}},
implementing an embedded Lisp intepreter for a small subset of this programming
language, and
\emph{html}\footnote{\href{http://hackage.haskell.org/package/html}{http://hackage.haskell.org/package/html}},
defining a combinator library for constructing HTML values.
%
These libraries also come with serialization functions to map Haskell values
into corresponding test case files.

We firstly compiled instrumented versions of the target programs in a way that
they also return the execution path followed in the source code every time we
run them with a given input test case.
%
This let us distinguish the amount of different execution paths that a randomly
generated corpus can trigger.
%
We then used the ADTs defined on the chosen libraries to derive random
generators using \dragen and \dragenp, including structural information
extracted from the library's codebase in the case of the latter.
%
Then, we proceeded to evaluate the code coverage triggered by independent,
randomly generated corpora of different sizes varying from 100 to 1000 test
cases each.
%
In order to remove any external bias, we derived generators optimized to follow
\emph{a uniform distribution of constructors (and pattern matchings or function
  calls in the case \dragenp), and carefully adjusted their generation sizes to
  match the average test case size in bytes}.
%
This way, any noticeable difference in the code coverage can be attributed to
the present (or lack thereof) structural information when generating the test
cases.
%
Additionally, to achieve statistical significance we repeated each experiment 30
times with independently generated sets of random test cases.

\begin{figure*}[t]
  \centering
  \input{tikz/lisp.tex}
  \hspace{5pt}%
  \input{tikz/html.tex}
  \caption{Path coverage comparison between \dragen (\ref{exp:dragen}) and
    \dragenp (\ref{exp:dragenp}). }
  \label{fig:coverage}
\end{figure*}
%
Fig. \ref{fig:coverage} illustrates the mean number of different execution paths
triggered for different sizes of the corpus size and derivation tool, including
error bars indicating the standard error of the mean on each case.
%
We proceed to describe each case study and our findings in detail as follows.

\subsection{Branching on input data}

In this first case study we wanted to evaluate the observed code coverage
differences when considering structural information present on functions pattern
matchings.

Our chosen library encodes Lisp S-expressions essentially as lists of symbols,
represented as plain strings; and literal values like booleans or integers.
%
In order to interpret Lisp programs, this unified representation of data and
code requires this library to pattern match against common patterns like
let-bindings, if-then-else expressions and arithmetic operators among others.
%
In particular, each one of these patterns match a against special symbol of the
Lisp syntax like |"let"|, |"if"| or |"+"|; and their corresponding
sub-expressions.
%
We extracted this structural information and included it into the generation
specification of our random Lisp values---which were generated by randomly
picking from a total of 6 data constructors and 8 different pattern matchings.
%
By doing this, we obtained a code coverage improvement of approximately $4\%$
using \dragenp with respect to the one obtained with \dragen (see Figure
\ref{fig:coverage} (a)).
%
While it seems a little improvement, it is worth noticing that we are generating
highly structured input, i.e., Lisp programs.
%
When testing interpreters, we argue that an improvement of $4\%$ is not
negligible considering the little effort that took us to specify the pattern
matchings.
%\todo[inline, author=AM]{Shall we say anything justifing this little
%  improvement?}

\subsection{Abstract interfaces}

@@

The library we used for this purpose represents HTML values very much in the
same way as we exemplify in Section \ref{sec:randomtesting}, i.e., defining a
small set of general constructions representing plain text and tags---although
this library also supports HTML tag attributes as well.
%
Then, this representation is extended with a large abstract interface consisting
of combinators representing common HTML tags and tag attributes---equivalent to
the combinators |div|, |bold| and |hr| illustrated in Section \ref{sec:sources}.

In this case study we included the structure information present on the abstract
interface of this library into the generation specification of random HTML
values, resulting in a generation process that randomly picked among 4 data
constructors and 163 abstract combinators.
%
With this large amount of additional structural information, we observed an
increase of up to $83\%$ in the code coverage obtained with \dragenp with
respect to the one observed with \dragen (see Figure \ref{fig:coverage} (b)).
%
A manual inspection of the corpora generated with each tool revealed us that, in
general terms, the test cases generated with \dragen rarely represent
syntactically correct HTML values, consisting to a large extent of random
strings within and between HTML tag delimeters (|"<"|, |">"| and |"/>"|).
%
On the other hand, test cases generated with \dragenp encode much more
interesting structural information, being mostly syntactically correct.
%
We found that, in many cases, the test cases generated with \dragenp were
parsed, analysed and reported as valid HTML values by the target application.

With these results we are confident that including the structural information
present on the user codebase improves the testing performance.
%
We consider that our approach is particularly useful when the data types
encoding the shape of our data are vague or not sufficiently structured to be
used to derive powerful random generators with the common derivation techniques
based only on ADT definitions.

### 2019-01-29 paper/synthesis.tex

Aliquam erat volutpat. Nunc eleifend leo vitae magna. In id erat non orci
commodo lobortis. Proin neque massa, cursus ut, gravida ut, lobortis eget,
lacus. Sed diam. Praesent fermentum tempor tellus. Nullam tempus. Mauris ac
felis vel velit tristique imperdiet. Donec at pede. Etiam vel neque nec dui
dignissim bibendum. Vivamus id enim. Phasellus neque orci, porta a, aliquet
quis, semper a, massa. Phasellus purus. Pellentesque tristique imperdiet tortor.
Nam euismod tellus id erat.

\begin{align*}
  % M_{|Html|} =
  \begin{tikzpicture}[baseline=-0.65ex]
    \matrix
    [ matrix of math nodes
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      M_{ADT\!\rightarrow\!ADT} \&
      M_{ADT\!\rightarrow\!Patterns} \&
      M_{ADT\!\rightarrow\!Interface} \\
      M_{Patterns\!\rightarrow\!ADT} \&
      M_{Patterns\!\rightarrow\!Patterns} \&
      M_{Patterns\!\rightarrow\!Interface} \\
      M_{Interface\!\rightarrow\!ADT} \&
      M_{Interface\!\rightarrow\!Patterns} \&
      M_{Interface\!\rightarrow\!Interface} \\
    };
  \end{tikzpicture}
\end{align*}

% \begin{align*}
%   M_C
% %   % =
% %   % \begin{bmatrix}
% %   %   m_{ij}  \\
% %   % \end{bmatrix}_{i, j \in \{|Leaf|, |NodeA|, |NodeB|\}}
% %   = \qquad\quad

% %   % \begin{tikzpicture}[baseline=-0.65ex]
% %   %   \matrix[
% %   %   matrix of math nodes,
% %   %   column sep=0ex,
% %   %   left delimiter={[},
% %   %   right delimiter={]}
% %   %   ] (m)
% %   %   {
% %   %     0               & 0                & 0\\
% %   %     2 \cdot p_{|Leaf|} & 2 \cdot p_{|NodeA|} &2 \cdot p_{|NodeB|}\\
% %   %      p_{|Leaf|}  &  p_{|NodeA|} &  p_{|NodeB|}\\
% %   %   };
% %   %   % \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle |Leaf|$};
% %   %   % \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle |NodeA|$};
% %   %   % \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle |NodeB|$};
% %   %   % \node[left,overlay] at ([xshift=-4.5ex]m-1-1.west) {$\scriptstyle |Leaf|$};
% %   %   % \node[left,overlay] at ([xshift=-1.5ex]m-2-1.west) {$\scriptstyle |NodeA|$};
% %   %   % \node[left,overlay] at ([xshift=-3.0ex]m-3-1.west) {$\scriptstyle |NodeB|$};
% %   % \end{tikzpicture}
% %   \label{eq:treepmatrix}
% \end{align*}

>>>>

Characterizing the distribution of values of an arbitrary random generator is a
hard task.
%
It requires to model every random choice that it could possibly make to generate
a value.
%
In a recent work \tocite, we have shown that it is possible to
\emph{analitically} predict the average distribution of data constructors
produced by a random generators automatically derived considering only ADT
definitions---like the one presented on Section \ref{sec:randomtesting}.
%
For this purpose, we found that random generation of ADT values can be
characterized using the theory of \emph{branching processes} \tocite.
%
This probabilistic theory was originally conceived to predict the growth and
extinction of royal family trees the Victorian Era, being later applied to a
wide variety of research areas.
%
In this work we extend this model to predict the average distribution of values
of random generators derived considering structural information coming from
pattern matchings and abstract interfaces.

Essentially, a branching process is a special kind of Markov process that models
the evolution of a population of \emph{individuals} of different kinds across
discrete time steps known as \emph{generations}.
%
Each kind of individual is expected to produce an average number of offspring of
(possibly) different kinds from one generation to the next one.
%
In our original setting, we shown that each different data constructor can be
considered as an individual on its own.
%
On the other hand, any ADT value can be seen as a tree where each node
represents a root data constructor and has its sub-expressions as
sub-trees---hence note the similarity with family trees.
%
In this light, each tree level of a generated value can be seen as a generation
of our generation model.

We note $G_n$ to the vector of real numbers that represents the average
distribution of constructors that a random generator produces in the $n$-th
generation, where each element of this vector represents each different data
constructor involved in the generation process.
%
Then, it is possible to predict expected number of constructors a the generation
$n$ by the means of the next formula:
%
\begin{align*}
  E[G_n]^T = E[G_0]^T \cdot M^n
\end{align*}

\begin{align*}
  E \left[ P_n \right]
  =
  \beta_{Html}^T \cdot \left( \frac{I - (M_{Html})^{n+1}}{I - M_{Html}}\right)
\end{align*}

\begin{align*}
  E[G_n]^T = \beta^T \cdot (\beta \cdot \mathcal{P})^{n+1}
\end{align*}

\newcommand{\ph}{\phantom{\square}}
\newcommand{\el}{\square}
\begin{equation*}
  M_{T} = \ \
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=-1pt
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (m)
    {
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Text
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Sing
      \ph \& \ph \& \el \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Tag
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % Join
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \el \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
      \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \& \ph \\ % simplify#1
    };
    % Color squares
    \fill[red, opacity=0.075]
    (m-1-1.north west)
    -|| ([xshift=-1pt]m-1-3.north east)
    -|| ([xshift=-1pt, yshift=1pt]m-3-3.south east)
    -|| ([xshift=1pt, yshift=1pt]m-3-1.south west)
    -|| cycle;
    \fill[blue, opacity=0.075]
    ([yshift=-1pt]m-4-1.north west)
    -|| ([xshift=1pt]m-1-3.north east)
    -|| ([xshift=0pt, yshift=-1pt]m-1-9.south east)
    -|| ([xshift=0pt, yshift=0pt]m-9-9.south east)
    -|| ([xshift=-1pt, yshift=0pt]m-9-1.south east)
    -|| cycle;
    % Upper labels
    \node[above,text depth=1pt] at (m-1-1.north) {$\scriptstyle C_1$};
    \node[above,text depth=1pt] at (m-1-2.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-3.north) {$\scriptstyle C_i$};
    \node[above,text depth=1pt] at (m-1-4.north) {$\scriptstyle P_1$};
    \node[above,text depth=1pt] at (m-1-5.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-6.north) {$\scriptstyle P_j$};
    \node[above,text depth=1pt] at (m-1-7.north) {$\scriptstyle F_1$};
    \node[above,text depth=1pt] at (m-1-8.north) {$\scriptstyle \cdots$};
    \node[above,text depth=1pt] at (m-1-9.north) {$\scriptstyle F_k$};
    % Left labels
    \node[left,overlay] at (m-1-1.west) {$\scriptstyle C_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-2-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-3-1.west) {$\scriptstyle C_i$};
    \node[left,overlay] at (m-4-1.west) {$\scriptstyle P_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-5-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-6-1.west) {$\scriptstyle P_j$};
    \node[left,overlay] at (m-7-1.west) {$\scriptstyle F_1$};
    \node[left,overlay] at ([xshift=-4pt, yshift=2pt]m-8-1.west) {$\scriptstyle \vdots$};
    \node[left,overlay] at (m-9-1.west) {$\scriptstyle F_k$};
    % Dashed lines
    \draw[densely dashed] ([xshift=-2pt]m-3-1.south west) -- ([xshift=2pt]m-3-9.south east);
    \draw[densely dashed] ([xshift=-2pt]m-6-1.south west) -- ([xshift=2pt]m-6-9.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-3.north east) -- ([yshift=-2pt]m-9-3.south east);
    \draw[densely dashed] ([yshift=2pt]m-1-6.north east) -- ([yshift=-2pt]m-9-6.south east);
    \draw[dotted, semithick] (m-3-1.west) -- ([xshift=3pt]m-3-3.west);
    \draw[dotted, semithick] (m-1-3.north) -- ([yshift=-3pt]m-3-3.north);
    \draw[dotted, semithick] (m-6-1.west) -- ([xshift=4pt]m-6-7.west);
    \draw[dotted, semithick] (m-1-7.north) -- ([yshift=-3pt]m-6-7.north);
  \end{tikzpicture}
\end{equation*}

\begin{equation*}
  \beta_{Html} = \qquad \qquad
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      0 \\ % Text
      0 \\ % Sing
      1 \\ % Tag
      2 \\ % Join
      0 \\ % simplify#1
      2 \\ % simplify#1
      0 \\ % hr
      1 \\ % div
      1 \\ % bold
    };
    \draw[densely dashed] ([xshift=-3pt]v-4-1.south west) -- ([xshift=3pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
    \node[left,overlay] at ([xshift=-2pt]v-1-1.west) {$\scriptstyle Text$};
    \node[left,overlay] at ([xshift=-2pt]v-2-1.west) {$\scriptstyle Single$};
    \node[left,overlay] at ([xshift=-2pt]v-3-1.west) {$\scriptstyle Tag$};
    \node[left,overlay] at ([xshift=-2pt]v-4-1.west) {$\scriptstyle Join$};
    \node[left,overlay] at ([xshift=-2pt]v-5-1.west) {$\scriptstyle simplify\#1$};
    \node[left,overlay] at ([xshift=-2pt]v-6-1.west) {$\scriptstyle simplify\#2$};
    \node[left,overlay] at ([xshift=-2pt]v-7-1.west) {$\scriptstyle hr$};
    \node[left,overlay] at ([xshift=-2pt]v-8-1.west) {$\scriptstyle div$};
    \node[left,overlay] at ([xshift=-2pt]v-9-1.west) {$\scriptstyle bold$};
  \end{tikzpicture}
  \hspace{20pt}
  \mathcal{P}_{Html} =
  \begin{tikzpicture}
    [ baseline=-0.65ex
    , every left delimiter/.style={xshift=.5em}
    , every right delimiter/.style={xshift=-.5em}
    ]
    \matrix
    [ matrix of math nodes
    , column sep=0ex
    , row sep=-1pt
    , ampersand replacement=\&
    , left delimiter={[}
    , right delimiter={]}
    ] (v)
    {
      p_{Text} \\
      p_{Single} \\
      p_{Tag}  \\
      p_{Join} \\
      p_{simplify\#1} \\
      p_{simplify\#2} \\
      p_{hr} \\
      p_{div} \\
      p_{bold} \\
    };
    \draw[densely dashed] ([xshift=-15pt]v-4-1.south west) -- ([xshift=15pt]v-4-1.south east);
    \draw[densely dashed] ([xshift=-3pt]v-6-1.south west) -- ([xshift=3pt]v-6-1.south east);
  \end{tikzpicture}
\end{equation*}

\paragraph{End-to-End Prediction}
%
% So far we have shown that our prection model is able to predict the distribution
% of data constructors, composite pattern values and interface function calls that
% our automatically derived random generators will produce on average.

It is possible to provide further ``end-to-end'' guarantees on the generated
values if we restrict the generation process to only bare data constructors and
pattern matching values.
%
In particular, we can stop considering pattern matching values as atomic
constructions and start seeing them as compositions of data constructors.
%
This way is possible to obtain the average ``total'' number of data constructors
that our generators will produce, regardless of if they are generated on their
own, or as part of a pattern matching value.
%
We note this number as $E^\downarrow\![\_]$.

The prediction only needs to consider the prediction of each data constructor
and add the average number of constructors that we will additionally generate
within pattern matching values.
%
For instance, we can calculate the total number of constructors |Text| and
|Join| that we will generate by expanding the average number of pattern values
$simplify\#1$ and $simplify\#2$ that we generate on average as follows:
%
{\small
  \begin{align*}
    E^\downarrow\!\left[ Text \right]
    = E\left[ Text \right]
    + 2\! \cdot\! E\left[ simplify\#1 \right]
    + E\left[ simplify\#2 \right] \\
    E^\downarrow\!\left[ Join \right]
    = E\left[ Join \right]
    + E\left[ simplify\#1 \right]
    + 2\! \cdot\! E\left[ simplify\#2 \right]
  \end{align*}
}
%
Where $simplify\#1$ carries two |Text| and one |Join| data constructors---and
viceversa for $simplify\#2$.

Note that this expansion cannot be applied if we also generate random values
containing abstract interface function calls, as we cannot predict the output of
an arbitrary function without solving the halting problem in the process.
